
##Reproducible Research: Peer Project 1

Author: Elva Yan
Date: Saturday, January 17, 2015



###Loading and preprocessing the data:

Frist of all,set the working directory to be where this R Markdown and all of the analysis is located. Then, load data into data frame named "activity":
```{r,echo=T}
# Remove all current data in workspace:
rm(list=ls())

library(knitr)
setwd("C:/Users/elva.yan/Documents/Training/Coursera/Reproductible Research/repdata-data-activity")
activity<- read.csv("activity.csv")
```



###What is the mean total number of steps taken per day?

(1) Plot a histogram of the total number of steps taken per day:
```{r,echo=T}
step1<- tapply(activity$steps, activity$date, FUN=sum, na.rm=TRUE)
hist(step1,col = "green",main = "Histogram of the Total Steps Taken Per Day",xlab = "Total Steps Taken Per Day", ylab="",breaks = 30)

```

(2) Calculate the mean and median total number of steps taken per day:

The mean value is:
```{r,echo=T}
mean(step1, na.rm=TRUE)
```

The median value is:
```{r,echo=T}
median(step1, na.rm=TRUE)
```



###What is the average daily activity pattern?

(1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r,echo=T}
library(ggplot2)
average1 <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),FUN=mean, na.rm=TRUE)
ggplot(data=average1, aes(x=interval, y=steps)) + geom_line() + xlab("Interval") + ylab("Average Steps Per Day")

```

(2) On average across all the days in the dataset, which 5-minute interval contains the maximum number of steps are:
```{r,echo=T}
average1[which.max(average1$steps),]
```



###Imputing missing values

Missing data needed to be imputed in this case. Asimple imputation approach was required: Missing values were imputed by inserting the average for each interval. 

(1) Firstly, calculate and the total number of missing values in the dataset and we can get:
```{r,echo=T}
NaValue <- is.na(activity$steps)
table(NaValue)
```
 
(2) Secondly, create a new dataset that is equal to the original dataset but with the missing data being filled in:
```{r,echo=T}
value <- function(steps, interval) {
    fill <- NA
    if (!is.na(steps))
        fill <- c(steps)
    else
        fill <- (average1[average1$interval==interval, "steps"])
    return(fill)
}
fData <- activity
fData$steps <- mapply(value, fData$steps, fData$interval)
```

(3) Then, make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day:
```{r,echo=T}
total<- tapply(fData$steps, fData$date, FUN=sum)
hist(total,col = "green",main = "Histogram of the Total Steps Taken Per Day",xlab = "Total Steps Taken Per Day", ylab="",breaks = 30)
mean(total)
median(total)
```

The new mean value is:
```{r,echo=T}
mean(total)
```

The new median value is:
```{r,echo=T}
median(total)
```



###Are there differences in activity patterns between weekdays and weekends?

Created a plot to compare and contrast number of steps between the week and weekend. There is a higher peak earlier on weekdays, and more overall activity on weekends: Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis):

```{r,echo=T}
daytype <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Saturday", "Sunday"))
        return("Weekend")
    else 
        return("Weekday")
}
fData$date <- as.Date(fData$date)
fData$day <- sapply(fData$date, FUN=daytype)

```

```{r,echo=T}
average2 <- aggregate(steps ~ interval + day, data=fData, mean)
library(lattice)
xyplot(steps~interval|day,average2, type = "l", layout = c(1,2), main = "Time Series Plot of Interval and Average Steps", xlab = "Interval", ylab = "Average Number of Steps Taken")

```
